An Extendable \proglang{Python} Implementation of Robust Optimisation Monte Carlo

### 2022-02-02 latex_files/ch3/3-0-design-principles.tex
Add comments and some changes. Ask HP.

For designing an extendable implementation of ROMC, we firstly define
ROMC as a sequence of algorithmic steps. The algorithmic view of ROMC
serves as the driver for the implementation at \pkg{ELFI}. At a high
level, ROMC can be split into the trainind the inference part. In
general terms, the training part covers all steps for estimating the
proposal regions and the inference part for getting weighted
samples. Algorithm~\ref{alg:romc_algorithm} defines ROMC formally;
Steps 2-10 (before the vertical line) form the training part, while
steps 12-17 the inference part.

\begin{algorithm}[!ht]
	\caption{ROMC}\label{alg:romc_algorithm}
	\begin{algorithmic}[1]
    \Procedure{ROMC}{$M_r(\thetab), n_1, n_2, \epsilon$}
    \For{$i \gets 1 \textrm{ to } n_1$}
      \State $\vb_i \sim p(\vb)$ \Comment{Draw nuisance variables}
      \State $d_i(\thetab) = M_d(\thetab, \vb=\vb_i)$ \Comment{Define distance function}
      \State $\thetab_i^* = \text{argmin}_{\thetab} d_i$, $d_i^*=d_i(\thetab_i^*)$ \Comment{Solve optimization problem}
      \If{$d^*_i > \epsilon$}
        \State Go to 2 \Comment{Filter solution}
      \EndIf
      \State Define $q_i$ over $\mathcal{\hat{S}}$ \Comment{Estimate proposal area}
      \State (Optional) Fit $\Tilde{d}_i$ on $\mathcal{\hat{S}}$ \Comment{Fit surrogate model}
      \\\hrulefill
      \For{$j \gets 1 \textrm{ to } n_2$}      
        \State $\thetab_{ij} \sim q_i$, compute $w_{ij}$ \Comment{Sample}
      \EndFor
    \EndFor
    \State $E_{p(\thetab|\data)}[h(\thetab)]$ \Comment{Estimate an expectation}
    \State $p_{d,\epsilon}(\thetab)$ \Comment{Evaluate the unnormalized posterior}
    \EndProcedure
	\end{algorithmic}
\end{algorithm}

@@

After obtaining the optimal points $\thetab^*$, the training part gets
completed by estimating the proposal
regions. Algorithm~\ref{alg:region_construction} describes the line
search approach for finding the region's boundaries. An important step
for the proposal region estimation is deciding the axes of the
bounding box i.e. the directions $\vb_d$ we follow for reaching the
boundaries. We approximate them as the direction of the highest
curvature of $d_i$ at $\thetab_i^*$. This estimation is given by the
eigenvalues of the Hessian matrix $\hessian_i$ of
$d_i$\footnote{Either the real distance $d_i$ or the Gaussian Process
  approximation $\hat{d}_i$} at $\thetab_i$. The Hessian matrix is
approximated numerically. In case where the distance function is the
Euclidean, the Hessian matrix can be also computed as
$\hessian_i = \jac_i^T\jac_i$, where $\jac_i$ is the Jacobian matrix
of the summary statistics $\Phi(M_d(\thetab, \vb=\vb_i))$ at
$\thetab_i^*$. The approximation through the Jacobian matrix has the
computational advantage of using only first-order derivatives.

\subsubsection*{Inference Part}
Performing the inference includes one or more of the following three
tasks; (a) sampling from the posterior
$ \thetab_i \sim p_{d, \epsilon}(\thetab|\data)$, (b) computing an
expectation $E_{\thetab|\data}[h(\thetab)]$ or (c) evaluating the
unnormalised posterior $p_{d, \epsilon}(\thetab|\data)$. Sampling is
performed by getting $n_2$ samples from each proposal distribution
$q_i$. For each sample $\thetab_{ij}$, the distance
function\footnote{As before, if a surrogate model $\hat{d}$ is
  available, it can be utilised as the distance function.} is
evaluated for checking if it lies inside the acceptance
region. Algorithm~\ref{alg:sampling_GB} defines the steps for
computing a weighted sample. Computing an expectation can be done
easily after weighted samples are obtained using the
equation~\ref{eq:expectation}, so we do not discuss it
separately. Evaluating the unnormalised posterior requires the
deterministic functions $g_i$ and the prior distribution
$p(\thetab)$\footnote{There is no need for solving the optimisation
  problems and building the proposal regions.}. The evaluation
requires iterating over all $g_i$ and evaluating the distance from the
observed data.

@@

\begin{algorithm}[!ht]
	\caption{Approximation of $\mathcal{\hat{S}}_i$ with a bounding box; Requires, a model of distance $d_i(\thetab)$, optimal point $\thetab_i^*$, number of refinements $K$, step size $\eta\_\text{start}$, maximum iterations $M$ and curvature matrix $\hessian_i$ ($\jac_i^T\jac_i $ or GP Hessian)}\label{alg:region_construction}
	\begin{algorithmic}[1]
	\State Compute eigenvectors $\vb_{d}$ of $\hess_i$ {\scriptsize ($d = 1,\ldots,||\thetab ||)$}
	\For{$d \gets 1 \textrm{ to } ||\thetab||$}
		\State $\Tilde{\thetab} \gets \thetab_i^*$ \label{algstep:box_constr_start}
		\State $k \gets 0$
		\State $\eta \gets \eta\_\text{start}$ \Comment{Initialize $\eta$}
		\Repeat
          \State $j \gets 0$
        	\Repeat
            \State $\Tilde{\thetab} \gets \Tilde{\thetab} + \eta \ \vb_{d}$ \Comment{Large step size $\eta$.}
            \State $j \gets j + 1$
        	\Until{$d(f_i(\Tilde{\thetab}), \data) > \epsilon$ or $j \geq M$} \Comment{Check distance or maximum iterations}
        	\State $\Tilde{\thetab} \gets \Tilde{\thetab} - \eta \ \vb_{d}$
        	\State $\eta \gets \eta/2$ \Comment{More accurate region boundary}
        	\State $k \gets k + 1$
      \Until $k = K$
      \If{$\Tilde{\thetab} = \thetab_i^*$} \Comment{Check if no step has been done}
        \State $\Tilde{\thetab} \gets \Tilde{\thetab} + \dfrac{\eta\_{\text{start}}}{2^K} \vb_d$ \Comment{Then, make the minimum step}
      \EndIf
    	\State Set final $\Tilde{\thetab}$ as region end point. \label{algstep:box_constr_end}
    	\State Repeat steps~\ref{algstep:box_constr_start}~-~\ref{algstep:box_constr_end} for $\mathbf{v}_{d} = - \mathbf{v}_{d}$
	\EndFor
	\State Fit a rectangular box around the region end points and define $q_i$ as uniform distribution
	\end{algorithmic}
\end{algorithm}

>>>>

For designing an extendable implementation of ROMC, we firstly define
ROMC as a sequence of algorithmic steps. The algorithmic view of ROMC
serves as the driver for the implementation at \pkg{ELFI}. At a high
level, ROMC can be split into the training and the inference part. In
general terms, the training part covers all steps for estimating the
proposal regions and the inference part calculates the weighted
samples. Algorithm~\ref{alg:romc_algorithm} defines ROMC formally;
Steps 2-10 (before the horizontal line) and steps 12-17 form the training 
part and the inference part, respectively. \textbf{HP comment Algorithm 1: I got a bit 
lost with $M_r(\thetab)$ (Line 1) and $M_d(\thetab, \vb=\vb_i)$ (Line 4). 
What is their relationship? Also, are Lines 13, 16 and 17 still under construction?
Maybe refer to Eq. (X)s in the Algorithm, if wanting to keep the Algorithm compact.}

\begin{algorithm}[!ht]
	\caption{ROMC}\label{alg:romc_algorithm}
	\begin{algorithmic}[1]
    \Procedure{ROMC}{$M_r(\thetab), n_1, n_2, \epsilon$}
    \For{$i \gets 1 \textrm{ to } n_1$}
      \State $\vb_i \sim p(\vb)$ \Comment{Draw nuisance variables}
      \State $d_i(\thetab) = M_d(\thetab, \vb=\vb_i)$ \Comment{Define distance function}
      \State $\thetab_i^* = \text{argmin}_{\thetab} d_i$, $d_i^*=d_i(\thetab_i^*)$ \Comment{Solve optimisation problem}
      \If{$d^*_i > \epsilon$}
        \State Go to 2 \Comment{Filter solution}
      \EndIf
      \State Define $q_i$ over $\mathcal{\hat{S}}$ \Comment{Estimate proposal area}
      \State (Optional) Fit $\Tilde{d}_i$ on $\mathcal{\hat{S}}$ \Comment{Fit surrogate model}
      \\\hrulefill
      \For{$j \gets 1 \textrm{ to } n_2$}      
        \State $\thetab_{ij} \sim q_i$, compute $w_{ij}$ \Comment{Sample}
      \EndFor
    \EndFor
    \State $E_{p(\thetab|\data)}[h(\thetab)]$ \Comment{Estimate an expectation}
    \State $p_{d,\epsilon}(\thetab)$ \Comment{Evaluate the unnormalised posterior}
    \EndProcedure
	\end{algorithmic}
\end{algorithm}

@@

After obtaining the optimal points $\thetab^*$, the training part gets
completed by estimating the proposal
regions. Algorithm~\ref{alg:region_construction} describes the line
search approach for finding the region's boundaries. An important step
for the proposal region estimation is deciding the axes of the
bounding box i.e. the directions $\vb_d$ we follow for reaching the
boundaries. We approximate them as the direction of the highest
curvature of $d_i$ at $\thetab_i^*$. This estimation is given by the
eigenvalues of the Hessian matrix $\hessian_i$ of
$d_i$\footnote{Either the real distance $d_i$ or the Gaussian Process
  approximation $\hat{d}_i$} at $\thetab_i$. The Hessian matrix is
approximated numerically. In case where the distance function is the
Euclidean, the Hessian matrix can be also computed as
$\hessian_i = \jac_i^T\jac_i$, where $\jac_i$ is the Jacobian matrix
of the summary statistics $\Phi(M_d(\thetab, \vb=\vb_i))$ \textbf{is 
$M_d(\cdot, \cdot)$ the deterministic simulator function, or deterministic 
evaluation of the distance function? If so, what is it's relationship to 
the summary statistics?} at
$\thetab_i^*$. The approximation through the Jacobian matrix has the
computational advantage of using only first-order derivatives.

\subsubsection*{Inference Part}
Performing the inference includes one or more of the following three
tasks; (a) sampling from the posterior
$ \thetab_i \sim p_{d, \epsilon}(\thetab|\data)$, (b) computing an
expectation $E_{\thetab|\data}[h(\thetab)]$ or (c) evaluating the
unnormalised posterior $p_{d, \epsilon}(\thetab|\data)$. Sampling is
performed by getting $n_2$ samples from each proposal distribution
$q_i$. For each sample $\thetab_{ij}$, the distance
function\footnote{As before, if a surrogate model $\hat{d}$ is
  available, it can be utilised as the distance function.} is
evaluated for checking if it lies inside the acceptance
region. Algorithm~\ref{alg:sampling_GB} defines the steps for
computing a weighted sample. Computing the expectation is easy after weighted 
samples are obtained using the equation~\ref{eq:expectation}, so we do not 
discuss it separately. Evaluating the unnormalised posterior requires the
deterministic functions $g_i$ and the prior distribution
$p(\thetab)$\footnote{There is no need for solving the optimisation
  problems and building the proposal regions.}. The evaluation
requires iterating over all $g_i$ and evaluating the distance from the
observed data. \textbf{HP comment: Algorithm 2 - introduce the notation $||\thetab||$, or even 
better, define the variable separately, instead of using the same notation 
that was used for the norm - could it be $N$ if $\thetab \in \mathbb{R}^N$. 
Note that Line 20 would lead to an infinite loop.}

@@

\begin{algorithm}[!ht]
	\caption{Approximation of $\mathcal{\hat{S}}_i$ with a bounding box; 
  Requires: a model of distance $d_i(\thetab)$, 
  an optimal point $\thetab_i^*$, 
  a number of refinements $K$, 
  a step size $\eta\_\text{start}$, 
  maximum iterations $M$ and 
  a curvature matrix $\hessian_i$ ($\jac_i^T\jac_i $ or GP Hessian)}\label{alg:region_construction}
	\begin{algorithmic}[1]
	\State Compute eigenvectors $\vb_{d}$ of $\hess_i$ {\scriptsize ($d = 1,\ldots,||\thetab ||)$}
	\For{$d \gets 1 \textrm{ to } ||\thetab||$}
		\State $\Tilde{\thetab} \gets \thetab_i^*$ \label{algstep:box_constr_start}
		\State $k \gets 0$
		\State $\eta \gets \eta\_\text{start}$ \Comment{Initialize $\eta$}
		\Repeat
          \State $j \gets 0$
        	\Repeat
            \State $\Tilde{\thetab} \gets \Tilde{\thetab} + \eta \ \vb_{d}$ \Comment{Large step size $\eta$.}
            \State $j \gets j + 1$
        	\Until{$d(f_i(\Tilde{\thetab}), \data) > \epsilon$ or $j \geq M$} \Comment{Check distance or maximum iterations}
        	\State $\Tilde{\thetab} \gets \Tilde{\thetab} - \eta \ \vb_{d}$
        	\State $\eta \gets \eta/2$ \Comment{More accurate region boundary}
        	\State $k \gets k + 1$
      \Until $k = K$
      \If{$\Tilde{\thetab} = \thetab_i^*$} \Comment{Check if no step has been done}
        \State $\Tilde{\thetab} \gets \Tilde{\thetab} + \dfrac{\eta\_{\text{start}}}{2^K} \vb_d$ \Comment{Then, make the minimum step}
      \EndIf
    	\State Set final $\Tilde{\thetab}$ as region end point. \label{algstep:box_constr_end}
    	\State Repeat steps~\ref{algstep:box_constr_start}~-~\ref{algstep:box_constr_end} for $\mathbf{v}_{d} = - \mathbf{v}_{d}$
	\EndFor
	\State Fit a rectangular box around the region end points and define $q_i$ as uniform distribution
	\end{algorithmic}
\end{algorithm}

### 2022-02-02 latex_files/ch3/3-1-implementation-principles.tex
Add comments and some changes. Ask HP.

The overview of our implementation is illustrated in
figure~\ref{fig:romc_overview}; one may interpret
figure~\ref{fig:romc_overview} as a depiction of the main class and
the ellipses are the main routines of the class. Following
\proglang{Python}'s naming principles, the methods starting with an
underscore (green ellipses) represent internal (private) functions,
whereas the rest (blue ellipses) are the public methods that the user
interacts with. In figure~\ref{fig:romc_overview}, it can be easily
observed that the implementation follows consistently the algorithmic
view presented in~\ref{alg:romc_algorithm}. The training part includes
all the steps until the computation of the proposal regions
i.e. sampling the nuisance variables, defining the optimization
problems, solving them, constructing the regions and fitting local
surrogate models. The inference part comprises of evaluating the
unnormalized posterior (and the normalized when is possible), sampling
and computing an expectation. We also provide some utilities for
inspecting the training process, such as plotting the histogram of the
final distances or visualizing the constructed bounding
boxes. Finally, in the evaluation part, we provide two methods for
evaluating the inference; (a) computing the Effective Sample Size
(ESS) of the samples and (b) measuring the divergence between the
approximate posterior the ground-truth, if the latter is
available.\footnote{Normally, the ground-truth posterior is not
  available; However, this functionality is useful in cases where the
  posterior can be computed numerically or with an alternative method
  (i.e.\ ABC Rejection Sampling), and we would like to measure the
  discrepancy between the two approximations.}

\subsubsection*{Parallelizing ROMC method}

ROMC has the significant advantage of being fully parallelizable. We
exploit this fact by implementing a parallel version of all the major
components of the method; (a) solving the optimization problems, (b)
constructing bounding box regions, (c) sampling and (d) evaluating the
posterior. We parallelize these processes using the built-in
\proglang{Python} package \pkg{multiprocessing}. The specific package
enables concurrency, using subprocesses instead of threads, for
side-stepping the Global Interpreter (GIL). For activating the
parallel version of the algorithm, the user simpy has to pass the
argument \code{parallelize=True} when instantiating \code{ROMC}.

@@

For illustrating the functionalities we will use the following running
example introduced by~\cite{Ikonomov2019},

>>>>

The overview of our implementation is illustrated in
Figure~\ref{fig:romc_overview}; one may interpret
Figure~\ref{fig:romc_overview} as a depiction of the main class and
the ellipses are the main routines of the class. Following
\proglang{Python}'s naming principles, the methods starting with an
underscore (green ellipses) represent internal (private) functions,
whereas the rest (blue ellipses) are the public methods that the user
interacts with. In Figure~\ref{fig:romc_overview}, it can be easily
observed that the implementation follows consistently Algorithm~\ref{alg:romc_algorithm}. The training part includes
all the steps until the computation of the proposal regions
i.e.~sampling the nuisance variables, defining the optimisation
problems, solving them, constructing the regions and fitting local
surrogate models. The inference part comprises of evaluating the
unnormalised posterior (and the normalised when is possible), sampling
and computing an expectation. We also provide some utilities for
inspecting the training process, such as plotting the histogram of the
final distances or visualising the constructed bounding
boxes. Finally, in the evaluation part, we provide two methods for
evaluating the inference; (a) computing the Effective Sample Size
(ESS) of the samples and (b) measuring the divergence between the
approximate posterior the ground-truth, if the latter is
available.\footnote{Normally, the ground-truth posterior is not
  available; However, this functionality is useful in cases where the
  posterior can be computed numerically or with an alternative method
  (i.e.\ ABC Rejection Sampling), and we would like to measure the
  discrepancy between the two approximations.}

\subsubsection*{Parallelising ROMC method}

ROMC has the significant advantage of being fully parallelisable. We
exploit this fact by implementing a parallel version of all the major
components of the method; (a) solving the optimisation problems, (b)
constructing bounding box regions, (c) sampling and (d) evaluating the
posterior. We parallelise these processes using the built-in
\proglang{Python} package \pkg{multiprocessing}. The specific package
enables concurrency, using subprocesses instead of threads, for
side-stepping the Global Interpreter (GIL). For activating the
parallel version of the algorithm, the user simpy has to pass the
argument \code{parallelize=True} when instantiating \code{ROMC}.

@@

For illustrating the functionalities we will use the following running
example introduced by~\cite{Ikonomov2019}, \textbf{HP comment: what is c in Eq. \eqref{eq:1D_example_eq_2}
and in python snippet? As it is, it will not run. }

### 2022-02-02 latex_files/ch4/4-1-training.tex
updated some images
Add comments and some changes. Ask HP.

\noindent
This method (a) draws the nuisance variables, (b) defines the
optimisation problems and (c) solves them using either a
gradient-based optimizer or Bayesian optimization. These tasks are
compleed sequentially, as shown in figure~\ref{fig:romc_overview}. The
definition of the optimisation problems is performed by drawing
\code{n1} integer numbers from a discrete uniform distribution
$u_i \sim \mathcal{U}\{1, 2^{32}-1\}$. Each integer $u_i$ is the seed
used in \pkg{ELFI}'s random simulator. The seed $u_i$ is the
computational analogous of the random variables $\vb_i$, that were
defined in the previous chapter; freezing the \code{seed} absorbs all
the randomness of the simulator transforming it to a deterministic
function. Setting the argument \code{use_bo=True}, chooses the
Bayesian Optimisation scheme for obtaining $\thetab_i^*$. In this
case, apart from obtaining the optimal points $\thetab_i^*$, we also
fit a Gaussian Process (GP) as surrogate model $\hat{d}_i$. In this
scenario, in the following steps, $\hat{d}_i(\thetab)$ replaces
$d_i(\thetab)$ as the distance function.

@@

\noindent
This function helps the user decide which threshold $\epsilon$ to
use. It plots a histogram of the distances at the optimal point
$d_i(\thetab_i^*) : \{i = 1, 2, \ldots, n_1\}$ or $\hat{d}_i^*$ in case
\code{use_bo=True}. The function accepts all keyword arguments and
forwards them to the underlying \code{matplotlib.hist()} function; in
this way the user may customise some properties of the histogram, such
as the number of bins or the range of values.

@@

\noindent
This method constructs a bounding box around the optimal point
$\thetab_i^* : i = 1, 2, \ldots, n_1$ following
Algorithm~\ref{alg:region_construction}. The Hessian matrix is
approximated based on the Jacobian
$\hess_i = \jac_i^T \jac_i$\footnote{The Jacobian matrix is computed
  on the output of the summary statistics
  $\nabla_{\thetab} \Phi(M_d(\thetab, \vb_i))$ at $\thetab_i^*$} if
gradients have been used, or as the Hessian of the GP, in case of
Bayesian Optimisation. The eigenvectors, which are the axes of the
bounding box, are computed using the function
\code{numpy.linalg.eig()}. Afterwards, the limits along the
eigenvectors are computed with
Algorithm~\ref{alg:region_construction}.

@@

\noindent
The function \code{fit_posterior} is a syntactic sugar for applying
\code{solve_problems} and \\ \code{estimate_regions} into a single
step. The function \code{visualize_region} can be used for plotting
the bounding box around the optimal point, when the parameter space is
$1D$ or $2D$. The argument \code{i} is the index of the corresponding
optimization problem i.e.\ $i<n_1$. \code{compute_eps} returns the
appropriate distance value $d_{i=\kappa}^*$ where
$\kappa = \lfloor \frac{quantile}{n} \rfloor$ from the collection
$\{ d_i^* \} \forall i = \{1, \ldots, n\}$ where $n$ is the number of
accepted solutions. It can be used to automate the selection of the
threshold $\epsilon$, e.g.\ \code{eps=romc.compute_eps(quantile=0.9)}.

@@

% \begin{figure}[h]
%   \begin{center}
%     \includegraphics[width=0.48\textwidth]{./latex_files/images/chapter3/example_theta_dist.png}
%     \includegraphics[width=0.48\textwidth]{./latex_files/images/chapter3/example_region_samples.png}
%     \end{center}
%     \caption[Histogram of distances at the 1D example.]{Histogram of
%       distances and visualisation of a specific region.}
%     \label{fig:example_training_hist}
% \end{figure}

% \begin{figure}[h]
%   \begin{center}
%     \includegraphics[width=0.32\textwidth]{./latex_files/images/chapter3/example_theta_dist.png}
%     \includegraphics[width=0.32\textwidth]{./latex_files/images/chapter3/example_region_samples.png}
%     \includegraphics[width=0.32\textwidth]{./latex_files/images/chapter3/example_posterior.png}

%     \end{center}
%     \caption[Histogram of distances at the 1D example.]{Histogram of
%       distances and visualisation of a specific region.}
%     \label{fig:example_training_hist}
%   \end{figure}

\begin{figure}[ht]
  \begin{center}    
    \input{latex_files/images/chapter3/example_gt.tex}
    \input{latex_files/images/chapter3/example_posterior.tex}
  \end{center}
  \caption[Histogram of distances at the 1D example.]{Histogram of
    distances and visualisation of a specific region.}
  \label{fig:running_example_romc_inference}
\end{figure}

>>>>

\noindent
This method (a) draws the nuisance variables, (b) defines the
optimisation problems and (c) solves them using either a
gradient-based optimiser or Bayesian optimisation. These tasks are
completed sequentially, as shown in Figure~\ref{fig:romc_overview}. The
definition of the optimisation problems is performed by drawing
\code{n1} integer numbers from a discrete uniform distribution
$u_i \sim \mathcal{U}\{1, 2^{32}-1\}$. Each integer $u_i$ is the seed
used in \pkg{ELFI}'s random simulator. The seeds $u_i$ are
computationally analogous to the random variables $\vb_i$, that were
defined in the previous chapter; setting the \code{seed} eliminates all randomness of the simulator transforming it to a deterministic
function. Setting the argument \code{use_bo=True}, chooses the
Bayesian Optimisation scheme for obtaining $\thetab_i^*$. In this
case, apart from obtaining the optimal points $\thetab_i^*$, we also
fit a GP as a surrogate model $\hat{d}_i$. In this
scenario, in the following steps, $\hat{d}_i(\thetab)$ replaces
$d_i(\thetab)$ as the distance function.

@@

\noindent
This function helps the user decide which threshold $\epsilon$ to
use. It plots a histogram of the distances at the optimal point
$d_i(\thetab_i^*) : \{i = 1, 2, \ldots, n_1\}$ or $\hat{d}_i^*$ in case
\code{use_bo=True}. The function accepts all keyword arguments and
forwards them to the underlying \code{matplotlib.pyplot.hist()} function; in
this way the user may customise some properties of the histogram, such
as the number of bins or the range of values.

@@

\noindent
This method constructs a bounding box around the optimal point
$\thetab_i^* : i = 1, 2, \ldots, n_1$ following
Algorithm~\ref{alg:region_construction}. The Hessian matrix is
approximated based on the Jacobian
$\hess_i = \jac_i^T \jac_i$\footnote{The Jacobian matrix is computed
  on the output of the summary statistics
  $\nabla_{\thetab} \Phi(M_d(\thetab, \vb_i))$ at $\thetab_i^*$} if
gradients have been used, or as the Hessian of the GP, in case of
Bayesian Optimisation. The eigenvectors are computed using the function
\code{numpy.linalg.eig()}, and used as the axes of the
bounding box. Afterwards, the limits along the
eigenvectors are computed with
Algorithm~\ref{alg:region_construction}.

@@

\noindent
The function \code{fit_posterior} is a syntactic sugar for applying
\code{solve_problems} and \\ \code{estimate_regions} into a single
step. The function \code{visualize_region} can be used for plotting
the bounding box around the optimal point, when the parameter space is
one- or two-dimensional. The argument \code{i} is the index of the corresponding
optimisation problem i.e.\ $i<n_1$. \code{compute_eps} returns the
appropriate distance value $d_{i=\kappa}^*$ where
$\kappa = \lfloor quantile \cdot n \rfloor$ from the collection
$\{ d_i^* \} \forall i = \{1, \ldots, n\}$ where $n$ is the number of
accepted solutions. It can be used to automate the selection of the
threshold $\epsilon$, e.g.\ \code{eps=romc.compute_eps(quantile=0.9)}.

@@

\begin{figure}[ht]
  \begin{center}    
    \input{latex_files/images/chapter3/example_gt.tex}
    \input{latex_files/images/chapter3/example_post.tex}
  \end{center}
  \caption[Histogram of distances at the one-dimensional example.]{Histogram of
    distances and visualisation of a specific region.}
  \label{fig:running_example_romc_inference}
\end{figure}

### 2022-02-02 latex_files/ch4/4-2-inference.tex
Add comments and some changes. Ask HP.

\noindent
This is the basic inference utility of the ROMC implementation; we
draw $n_2$ samples for each bounding box region. This gives a total of
$k \times n_2$, where $k < n_1$ is the number of the optimal points
remained after filtering\footnote{From the $n_1$ optimisation
  problems, only the ones with $d_i(\thetab_*) < \epsilon$ are
  maintained for building a bounding box}. The samples are drawn from
a uniform distribution $q_i$ defined over the corresponding bounding
box and the weight $w_i$ is computed as in
algorithm~\ref{alg:sampling_GB}.

@@

\noindent
This function computes the expectation
$E_{p(\thetab|\data)}[h(\thetab)]$ using
expression~\eqref{eq:expectation}. The argument \code{h} can be any
python \code{Callable}. \code{eval_unnorm_posterior} computes the
unnormalised posterior approximation using
expression~\eqref{eq:approx_posterior}. \code{eval_posterior}
evaluates the normalised posterior estimating the partition function
$Z = \int p_{d,\epsilon}(\thetab|\data)d\thetab$; with Riemann's
integral approximation. The approximation is computationally feasible
only in low-dimensional parametric spaces.

@@

In the following code snippet, we perform weighted sampling from the
ROMC approximate posterior. Afterwards, we use some of ELFI's
built-in tools to get a summary of the obtained samples. In figure
\ref{fig:running_example_romc_inference}, we observe the histogram of the weighted
samples and the acceptance region of the first deterministic function
(as before) alongside the samples obtained from it. Finally, in the
code snippet, we demonstrate how to use the \code{compute_expectation}
function; in the current example, we define \code{h} to compute
firstly the empirical mean and afterwards the empirical variance. In
both cases, the empirical result is close to the ground truth
$\mu = 0$ and $\sigma^2 = 1$. The
\code{romc.eval_unnorm_posterior(theta)} evaluates the posterior at
point $\theta$ using expression \eqref{eq:approx_posterior_omc}. The
\code{romc.eval_posterior(theta)} approximates the partition function
$Z = \int_{\thetab} p_{d,\epsilon}(\thetab|\data) d\thetab$ using the
Riemann approximation as explained above. In our simple example, this
utility can provide a nice plot of the approximate posterior as
illustrated in figure~\ref{fig:running_example_romc_inference}. We observe that the
approximation is quite close to the ground-truth posterior.

>>>>

\noindent
This is the basic inference utility of the ROMC implementation; we
draw $n_2$ samples for each bounding box region. This gives a total of
$k \times n_2$, where $k < n_1$ is the number of the optimal points
remained after filtering\footnote{From the $n_1$ optimisation
  problems, only the ones with $d_i(\thetab_*) < \epsilon$ are
  maintained for building a bounding box}. The samples are drawn from
a uniform distribution $q_i$ defined over the corresponding bounding
box and the weight $w_i$ is computed as in
Algorithm~\ref{alg:sampling_GB}.

@@

\noindent
This function computes the expectation
$E_{p(\thetab|\data)}[h(\thetab)]$ using
expression~\eqref{eq:expectation}. The argument \code{h} can be any
python \code{Callable}. Method \code{eval_unnorm_posterior} computes the
unnormalised posterior approximation using
expression~\eqref{eq:approx_posterior}. Method \code{eval_posterior}
evaluates the normalised posterior estimating the partition function
$Z = \int p_{d,\epsilon}(\thetab|\data)d\thetab$; with Riemann's
integral approximation. The approximation is computationally feasible
only in low-dimensional parametric spaces.

@@

In the following code snippet, we perform weighted sampling from the
ROMC approximate posterior. Afterwards, we use some of \pkg{ELFI}'s
built-in tools to get a summary of the obtained samples. In Figure
\ref{fig:running_example_romc_inference}, we observe the histogram of the weighted
samples and the acceptance region of the first deterministic function
(as before) alongside the samples obtained from it. Finally, in the
code snippet, we demonstrate how to use the \code{compute_expectation}
function; in the current example, we define \code{h} to compute
firstly the empirical mean and afterwards the empirical variance. In
both cases, the empirical result is close to the ground truth
$\mu = 0$ and $\sigma^2 = 1$. The
\code{romc.eval_unnorm_posterior(theta)} evaluates the posterior at
point $\theta$ using expression \eqref{eq:approx_posterior_omc}. The
\code{romc.eval_posterior(theta)} approximates the partition function
$Z = \int_{\thetab} p_{d,\epsilon}(\thetab|\data) d\thetab$ using the
Riemann approximation as explained above. In our simple example, this
utility can provide a nice plot of the approximate posterior as
illustrated in Figure~\ref{fig:running_example_romc_inference}. We observe that the
approximation is quite close to the ground-truth posterior.

### 2022-02-02 latex_files/ch4/4-3-evaluation.tex
Add comments and some changes. Ask HP.

\noindent
This function computes the divergence between the ROMC approximation
and the ground truth posterior. Since the computation is performed
using the Riemann's approximation, this method can only work in low
dimensional parameter spaces; it is suggested to be used for up to the
three-dimensional parameter space. As mentioned at the beginning of
this chapter, in a real-case scenario it is not expected the
ground-truth posterior to be available. However, in cases where the
posterior can be approximated decently well with a numerical approach
(as in the current example) or with some other inference approach,
this function can provide a numerical measure of the agreement between
the two approximations. The argument \code{step} defines the step used
in the Riemann's approximation and the argument \code{distance} can
take either the \code{Jensen-Shannon} or the \code{KL-divergence}
value, for computing the appropriate distance. \code{compute_ess}
computes the Effective Sample Size (ESS) using the following
expression~\cite{Sudman1967},

>>>>

\noindent
This function computes the divergence between the ROMC approximation
and the ground truth posterior. Because the computation is performed
using the Riemann's approximation, this method can only work in low
dimensional parameter spaces; it is suggested to be used for up to the
three-dimensional parameter space. As mentioned at the beginning of
this chapter, in a real-case scenario it is not expected the
ground-truth posterior to be available. However, in cases where the
posterior can be approximated acceptably well with a numerical approach
(as in the current example) or with some other inference approach,
this function can provide a numerical measure of the agreement between
the two approximations. The argument \code{step} defines the step used
in the Riemann's approximation and the argument \code{distance} can
be either \code{"Jensen-Shannon"} or \code{"KL-divergence"}, for computing 
the appropriate distance. Method \code{compute_ess}
computes the Effective Sample Size (ESS) using the following
expression~\cite{Sudman1967},
